---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r warning=FALSE, message=FALSE}
# Load tidyverse
library(tidyverse)
```

```{r}
localfile <- "activity.csv"
if (!file.exists(localfile)) {
  unzip ("activity.zip")
}

data <- read.csv(localfile)
data$date <- as.Date(data$date, "%Y-%m-%d")
summary(data)
```

## What is mean total number of steps taken per day?

```{r, fig.align='center'}
data %>% 
  filter (!is.na(steps)) %>% 
  group_by(date) %>% 
  summarise_all(sum) %>% 
  select(steps) %>% 
  unlist() %>% hist( main="Histogram of Total Steps per Day",
                     xlab="Steps per Day")
```
Mean and Median of the total number of steps taken per day:
```{r}
data %>% 
  filter (!is.na(steps)) %>% 
  group_by(date) %>% 
  summarise_all(sum) %>% 
  select(steps) %>% summarise(mean(steps),median(steps))
```


## What is the average daily activity pattern?

```{r, fig.align='center'}
b0 <- data %>% 
    select (interval, steps) %>%
    filter (!is.na(steps)) %>% 
    group_by(interval) %>% 
    summarise_all(mean) 

b0 %>%
    plot(type="l", main="Daily activity pattern",
           ylab="Average Steps")
    abline(v = b0$interval[b0$steps == max(b0$steps)], col="red")


```
Maximum numbers of steps is `r b0$interval[b0$steps == max(b0$steps)]`


## Imputing missingdata

```{r}
# Summary of Data
summary(data)
```

`r sum(is.na(data$steps))` steps data are missing


```{r,  fig.align='center'}
c0 <- data %>% 
    select (interval, date, steps) %>%
      mutate (observations = 1, missing = is.na(steps)) %>%
      group_by(date) %>%
      summarise_all(sum)

  

c0 %>% select (date, missing) %>%
    plot(type="b", main="Missing Data",
           ylab="Numbers of Missing Data")

```
Imputing Strategy: rounded mean of steps per interval
```{r, fig.align='center'}
# mean of steps per interval is already calculated and stored as b0
# for imputing missing data, rounded mean of steps per interval is used as numbers of steps are defined as integer

data_imputed <- data %>%
    rowwise() %>%
    mutate (steps2 = ifelse(is.na(steps), round(b0$steps[b0$interval==interval],0),steps), .after=steps) %>%
    mutate (delta_steps = steps - steps2, .after=steps2)

# check
d0 <- data_imputed %>% 
    select (interval, steps, steps2) %>%
    group_by(interval) %>% 
    summarise_all(mean, na.rm = TRUE)
plot(d0$steps,d0$steps2)
```
```{r}
#generating new data set
data_new <- data_imputed %>%
            select (steps = steps2, date, interval)
```
 
```{r, fig.align='center'}
data_new %>% 
 select (date,steps) %>%
  group_by(date) %>% 
  summarise_all(sum) %>% 
  select(steps) %>% 
  unlist() %>% hist( main="Histogram of Total Steps per Day with imputed data",
                     xlab="Steps per Day")
```



## Are there differences in activity patterns between weekdays and weekends?
```{r}

## new factor "daytype"
data_new <- data_new %>%
  mutate (weekday = weekdays(date)) %>%
  mutate (daytype= ifelse (as.POSIXlt(date)$wday %in% 1:5, "Weekday", "Weekend"))
data_new$daytype = as.factor(data_new$daytype)


data_new %>%
    select (interval, steps, daytype) %>%
    group_by(interval,daytype) %>% 
    summarise_all(mean) %>%
  
  ggplot() +
  aes (x=interval, y=steps, group=daytype) +
  geom_line() +
  facet_grid(daytype~ .)
  
```


